---
title: "R Notebook"
subtitle: "Data 608 - Module 1"
author: "Juan Falck"
output:
  html_document: default
  html_notebook: default
editor_options: 
  chunk_output_type: console
---

**Principles of Data Visualization and Introduction to ggplot2**

I have provided you with data about the 5,000 fastest growing companies in the US, as compiled by Inc. magazine. lets read this in:

```{r}

inc <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module1/Data/inc5000_data.csv", header= TRUE)

```

And lets preview this data:

```{r}

head(inc)

```


```{r}

summary(inc)

```

Think a bit on what these summaries mean. Use the space below to add some more relevant non-visual exploratory information you think helps you understand this data:

```{r, warning=FALSE, message=FALSE}
# Insert your code here, create more chunks as necessary
library(tidyverse)
library(scales)

#First let's glimpse the dataframe to get some insigts of all features (columns)
glimpse(inc)

# A Table for number of companies by State
inc %>%
  group_by(State) %>%
  drop_na(Employees) %>%
  summarise(total = n())

# A Table for number of Employees by State
inc %>%
  group_by(State) %>%
  drop_na(Employees) %>%
  summarise(total_employees = sum(Employees))

# A Table for Total Revenue by State in Millions
inc %>%
  group_by(State) %>%
  drop_na(Employees) %>%
  summarise(total_revenue_MM = sum(Revenue)/1000000)
  #mutate(total_revenue = comma_format(accuracy = 12)(total_revenue))

```

## Question 1

Create a graph that shows the distribution of companies in the dataset by State (ie how many are in each state). There are a lot of States, so consider which axis you should use. This visualization is ultimately going to be consumed on a 'portrait' oriented screen (ie taller than wide), which should further guide your layout choices.

```{r}
# Answer Question 1 here

inc %>%
  ggplot(aes(fct_rev(fct_infreq(State)))) +
  geom_bar(fill="#FF5733") +
  coord_flip() +
  theme_minimal() +
  labs(y="Companies within State", x="US State",
       title="Distribution of Companies by State")

```

## Question 2

Lets dig in on the state with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries. Create a plot that shows the average and/or median employment by industry for companies in this state (only use cases with full data, use R's `complete.cases()` function.) In addition to this, your graph should show how variable the ranges are, and you should deal with outliers.

```{r}
# Answer Question 2 here

# Lets see a table first to see totals by Industry
inc %>%
    filter(State=="NY") %>%
    filter(complete.cases(.)) %>%
    group_by(Industry) %>%
    summarise(total=n())

# now lets use Bloxplot in GGPLOT
inc %>%
  filter(State=="NY") %>%
  filter(complete.cases(.)) %>%
  ggplot(aes(x=reorder(Industry, Employees, FUN=median), y=Employees)) +
  geom_boxplot(fill="#FF5733") +
  scale_y_log10() +
  coord_flip() +
  theme_minimal() +
  labs(y="Number of Employees", x="Industry",
       title="Distribution of Employess by Industry in NY")


```

## Question 3

Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart that makes this information clear. Once again, the distribution per industry should be shown.

```{r}
# Answer Question 3 here

inc %>%
  filter(complete.cases(.)) %>%
  mutate(prod_employee = Revenue / Employees) %>%
  ggplot(aes(x= reorder(Industry, prod_employee, FUN = median), y=prod_employee)) +
  geom_boxplot(fill="#FF5733") +
  scale_y_log10(labels = label_comma()) +
  coord_flip() +
  theme_minimal() +
  labs(y="Revenue per Employee", x="Industry",
       title="Distribution Revenue by Employee by Industry")

```


